### 2024-05-16 paper/header.tex
Initial state

% **************GENERATED FILE, DO NOT EDIT**************

\title{Simulation of fracture and damage with Peridynamics.jl}

\author[1]{Kai Partmann}
\author[1]{Manuel Dienst}
\author[1]{Kerstin Weinberg}
\affil[1]{Chair of Solid Mechanics, University of Siegen, Siegen, Germany}

\keywords{Julia, Peridynamics, Fracture Mechanics, Simulations, Simulation Framework}

\hypersetup{
pdftitle = {Simulation of fracture and damage with Peridynamics.jl},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {Kai Partmann, Manuel Dienst, Kerstin Weinberg},
pdfkeywords = {Julia, Peridynamics, Fracture Mechanics, Simulations, Simulation Framework},
}