### 2024-05-16 paper/paper.tex
Add comments with infos from JOSS docs

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\section{Summary}
% from the JOSS docs:
% The paper should include a summary describing the high-level functionality and purpose of the software for a diverse, non-specialist audience.

\cite{bezanson2017julia}

\section{Statement of need}
% from the JOSS docs:
% The paper should include a Statement of need section that clearly illustrates the research purpose of the software and places it in the context of related work.

\section{Acknowledgments}
The authors gratefully acknowledge the support of the Deutsche Forschungsgemeinschaft (DFG) in the project \mbox{WE~2525/15-1}.

The authors gratefully acknowledge the support and the supercomputing resources of the Paderborn Center for Parallel Computing (https://pc2.uni-paderborn.de) and the University of Siegen (https://cluster.uni-siegen.de) used to conduct the research reported in this paper.

%\vadjust{\vfill\pagebreak}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template