### 2024-05-16 paper/paper.tex
first draft of begin of summary

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

%%---- definition of custom commands
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\vb}{\boldsymbol}
%%----

\begin{document}

\input{header}

\maketitle

\section{Summary}

\texttt{Peridynamics.jl} is an open source implementation of the nonlocal continuum mechanics formulation by Silling (ref?).
In peridynamics, a body is described by material points and, based on interactions between them, internal force densities are calculated. 
The equation of motion reads
\begin{equation}
    \varrho \, \vb{\ddot{u}}(\vb{X},t) = \vb{b}^{\mathrm{int}}(\vb{X},t) + \vb{b}^{\mathrm{ext}}(\vb{X},t) \qquad \forall \vb{X} \in \mathcal{B}_0, \; t \geq 0 \; ,
\end{equation} 
with the mass density $\varrho$, the point acceleration vector $\vb{\ddot{u}}$, and the point force density vectors $\vb{b}^{\mathrm{int}}$ and $\vb{b}^{\mathrm{ext}}$. 
In the original bond-based formulation (BB), the internal force density is calculated by
\begin{equation}
    \vb{b}^{\mathrm{int}}(\vb{X},t) = \int_\mathcal{H} \vb{f} \; \mathrm{d}V' \; ,
\end{equation}
with the \emph{pairwise force function} $\vb{f}$, which is evaluated for each material point $\vb{X}^i$ and depends on the deformation of the bonds to each point in its neighborhood $\mathcal{H}$.
These point families contain each point within a distance $\delta$, also called horizon.

To overcome the restrictions of the BB formulation with only one determinable material parameter, the state-based formulation was established.
Here not only the bonds to neighboring material points, but also the states of neighboring points influence the internal force density $\vb{b}^{\mathrm{int}}$ of each material point.
The equation reads
\begin{equation}
    \vb{b}^{\mathrm{int}} (\vb{X},t) = \int_\mathcal{H} \vb{t} - \vb{t}' \; \mathrm{d}V' \; ,
\end{equation}
with the force vector states $\vb{t}=\vb{t}(\vb{\Delta X}, t)$ and $\vb{t}'=\vb{t}(-\vb{\Delta X}, t)$.

The \texttt{Peridynamics.jl} package enables users to perform peridynamic simulations

\cite{bezanson2017julia}

\section{Statement of need}

\section{Acknowledgments}
The authors gratefully acknowledge the support of the Deutsche Forschungsgemeinschaft (DFG) in the project \mbox{WE~2525/15-1}.

The authors gratefully acknowledge the support and the supercomputing resources of the Paderborn Center for Parallel Computing (https://pc2.uni-paderborn.de) and the University of Siegen (https://cluster.uni-siegen.de) used to conduct the research reported in this paper.

%\vadjust{\vfill\pagebreak}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template